High-Performance Primitives for Distributed Memory Architectures using Asynchronous Events

### 2012-11-12 doc/llr-pldi2013/interface.tex
more minor fixes

If the only operations a task will perform on a region are reductions, a special reduction-only
instance can be created (line 55).  Rather than applying each reduction separately to the target instance,
the reduction operations can be accumulated in the reduction instance, and applied in bulk to 
the target using the {\em reduce\_to} operation (line 64).  This has several benefits.  First, it results
in much better performance, as we show in Section~\ref{subsec:reducmicro}, by reducing both
the total inter-node traffic and the overhead of that traffic while still allowing parallelism.
Second, a reduction-only instance is often smaller that the target instance, potentially improving
cache performance and helping address capacity issues for smaller memories (e.g. a GPU's device
memory).  For example, if reductions are being made to a single field in a larger structure, the 
reduction-only instance only needs to store that field and not the other fields that won't be 
modified by the reduction operation.

@@

An example using reduction-only regions is given in Figure~\ref{fig:reducevents} along with the resulting event graph.
Two reduction-only regions, {\tt inst1} and {\tt inst2}, are created in different memories to
be used with a reduction operation (lines 7-8).  (Reduction operations are registered at startup with integer IDs
as function pointers may not be portable across nodes.)  A physical region of the same class {\tt inst3}
is also created in a different memory (line 9).  Two tasks are launched that will apply
reductions locally into their respective reduction buffers (lines 10-11).  The resulting
events are then used to chain reduction copies back to {\tt inst3} after the tasks complete (lines 12-13).
The reduction operations accumulated in each of {\tt inst1} and {\tt inst2} are applied atomically
to {\tt inst3}.  The order in which the reduction operations are performed will be different from
the original order, but the requirement that reduction operations be commutative ensures the final
result is correct.  Finally, a merge on the events from the two reduction copies is 
performed and a final task using the results in {\tt inst3} is launched (lines 14-15).

>>>>

If the only operations a task will perform on a instance are reductions, a special reduction-only
instance can be created (line 55).  Rather than applying each reduction separately to the target instance,
the reduction operations can be accumulated in the reduction instance.  Using the {\tt reduce\_to} 
method (line 64), these accumulated reductions can be applied in bulk to another instance, which may be either
a normal instance or another reduction-only instance.  First, it results
in much better performance, as we show in Section~\ref{subsec:reducmicro}, by reducing both
the total inter-node traffic and the overhead of that traffic while still allowing parallelism.
Second, a reduction-only instance is often smaller than a normal instance, potentially improving
cache performance and helping address capacity issues for smaller memories (e.g. a GPU's device
memory).  For example, if reductions are being made to a single field in a larger structure, the 
reduction-only instance only needs to store that field and not the other fields that won't be 
modified by the reduction operation.

@@

An example using reduction-only instances is given in Figure~\ref{fig:reducevents} along with the resulting event graph.
Two reduction-only instances, {\tt inst1} and {\tt inst2}, are created in different memories to
be used with a reduction operation (lines 7-8).  (Reduction operations are registered at startup with integer IDs
as function pointers may not be portable across nodes.)  A normal instance {\tt inst3}
is also created in a different memory (line 9).  Two tasks are launched that will apply
reductions locally into their respective reduction instances (lines 10-11).  The resulting
events are then used to chain reduction copies back to {\tt inst3} after the tasks complete (lines 12-13).
The reduction operations accumulated in each of {\tt inst1} and {\tt inst2} are applied atomically
to {\tt inst3}.  The order in which the reduction operations are performed will be different from
the original order, but the requirement that reduction operations be commutative ensures the final
result is correct.  Finally, a merge on the events from the two reduction copies is 
performed and a final task using the results in {\tt inst3} is launched (lines 14-15).

### 2012-11-12 doc/llr-pldi2013/micro.tex
more minor fixes

For our experiment, there are 8 reduction tasks created for each node, one per processor.  Each task performs 4M reductions
into a histogram with 256K buckets.  The results for a variety of node counts are shown in Figure~\ref{fig:reducbench}.
They show that the version using reduction-only instance performs 
best and scales nearly linearly with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  The localized instance approach works well for a single node, 
but its inherently serial nature doesn't benefit from increasing node counts.  Finally, the version using individual
reduction messages is scalable, but the per-message overhead makes it nearly 1000x slower than the versions that are
able to move data in bulk.

>>>>

For our experiment, there are 8 reduction tasks created for each node, one per processor.  Each task performs 4M reductions
into a histogram with 256K buckets.  The input data is random, so although the average number of reductions per bucket is 16, there will be some buckets with more and some with fewer.  The results for a variety of node counts are shown in Figure~\ref{fig:reducbench}.
They show that the version using reduction-only instance performs 
best and scales nearly linearly with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  The localized instance approach works well for a single node, 
but its inherently serial nature doesn't benefit from increasing node counts.  Finally, the version using individual
reduction messages is scalable, but the need to send each reduction separately and the per-message overhead make it nearly 1000x slower than the reduction instance version that is
able to accumulate multiple reductions to the same location locally and move data in bulk.
